---
title: "Vertebra Morphometrics Project"
author: "John McPhail"
date: "2025-12-05"
output: html_document
---

## Overview

This project automates a geometric morphometric workflow for comparing vertebra
shape between two fish populations (Matnog vs Sacol) using landmark coordinates
stored in TPS files.

The analysis pipeline lives in `scripts/run_morphometrics.R` and is designed to
run without RStudio.

## Run the pipeline

```{r eval=FALSE}
system("Rscript ../scripts/run_morphometrics.R")
```

After running, outputs are written to:

- `results/summary.txt`
- `results/figures/`
- `data/processed/`

## Results (if present)

```{r}
summary_path <- file.path("..", "results", "summary.txt")
if (file.exists(summary_path)) {
  cat(readLines(summary_path), sep = "\n")
} else {
  cat("Run `Rscript scripts/run_morphometrics.R` from the repo root to generate outputs.")
}
```

```{r, echo=FALSE}
fig_dir <- file.path("..", "results", "figures")
figs <- c(
  "pca_pc1_pc2.png",
  "mean_shape_raw.png",
  "within_group_variation.png",
  "mean_shape_size_corrected.png"
)

paths <- file.path(fig_dir, figs)
paths <- paths[file.exists(paths)]

if (length(paths) > 0) {
  knitr::include_graphics(paths)
} else {
  cat("No figures found in `results/figures/`. Run the pipeline to generate them.")
}
```
